---
title: "Report2"
author: "Emily Wilhelm"
date: "2025-11-06"
output: html_document
---

```{r, here-i-am, include=FALSE}
here::i_am(
  "Report2.Rmd"
)
```

```{r, packages, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))
library(readr)
library(tidyverse)
library(lubridate)
library(gt)
install.packages("gtExtras")
library(gtExtras)
library(ggplot2)
```

## Georgia Air Quality Dataset 

This dataset includes air quality data that was collected every three days in Atlanta, GA throughout the year 2024, starting on 01/01/2024. The air pollution data was sourced from the EPA's Air Quality System. There are a total of 9 air quality monitoring sites in Atlanta which collect PM2.5 data. The some of the variables that comprise the dataset are the `Date`, the daily mean PM2.5 concentrations (ug/m3 LC) - `Daily.Mean.PM2.5.Concentrations`, the daily air qualtiy index (AQI) value - `Daily.AQI.Value`, air monitor site ID - `Site.ID`, site name - `Local.Site.Name`, and `County`. 

`Daily.AQI.Value` reports the overall air quality on a scale of 0-500. 

* 0-50 AQI: Good - air pollution poses little to no risk 

* 50-100 AQI: Moderate - some risk for thosee with high sensitivity to air pollution 

* 100-150 AQI: Unhealthy for sensitive groups - those with high sensitivity to air pollution may experience health effects

* 150-200 AQI: Unhealthy - some risk to the general public and risk of serious health effects for those with high sensitivity to air pollution 

* 200-300 AQI: Very unhealthy - increased health risks for everyone 

* 300-500 AQI: Hazardous - emergency conditions likely to affect everyone 

The air monitoring site names are: 

* Forest Park, Kennesaw, South Dekalb, NR-285, Fire Station #8, United Avenue, NR-GA Tech, Gwinnett, Mcdonough 


## Analysis Objectives 

* View monthly average PM2.5 (µg/m³) trends at each air monitoring site location using a table.   

* Compare daily mean PM2.5 (µg/m³) ranges each month between two air monitoring sites in the same county using box plots to see how much PM2.5 could vary in one county. 


```{r, load-files-needed, echo=FALSE}
Average_PM2.5 <- readRDS(
  here::here("output/Average_PM2.5.rds")
)

Fulton_data <- readRDS(
  here::here("output/Fulton_data.rds")
)
```

## Monthly Average PM2.5 (µg/m³) For Each Air Monitoring Site Location (2024)

```{r, table, echo=FALSE}
Average_PM2.5 %>% 
  gt() %>% 
  fmt_number(
    columns = -Month,  
    decimals = 2) %>% 
  gt_theme_pff()
```

## Plotting and Comparing Fulton County Air Monitoring Site Locations

```{r, plot, echo = FALSE}
Fulton_1 <- "NR-GA Tech"
Fulton_2 <- "United Avenue"

ggplot(Fulton_data, aes(x = Local.Site.Name, y = Daily.Mean.PM2.5.Concentration, fill = Local.Site.Name)) +
  geom_boxplot() + 
  facet_wrap(~Month) +
  labs(
    title = "Comparing Daily PM2.5 Range each Month Between Two Fulton County Air Monitoring Locations",
    x = "Month", 
    y = "Daily PM2.5 (µg/m³)",
    fill = "Local.Site.Name"
  ) +
  theme(
    axis.text.x = element_blank())
```

